---
title: "Second Assessment"
author: "20_Nishchal_Danuwar"
date: "2024-05-31"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Qestion number: 7
Loading necessary packages to work with air-quality dataset of R.
```{r, warning = FALSE}
library(dplyr)   # For data manipulation
library(tidyr)   # For data reshaping
library(ggplot2) # For visualization
library(broom)   # For tidy model outputs
```

a) Performing goodness-of-fit test on Temp variable to check if it follows normal distribution or not. 
Loading **airquality** data:
```{r}
data("airquality")
str(airquality)
head(airquality)
```

Now performing **goodness-of-fit** test on **Temp** variable using Shapiro-Wilk test for normality:
```{r}
shapiro_test <- airquality %>%
  summarise(p_value = shapiro.test(Temp)$p.value)

# Display the test results
shapiro_test
```
With p-value 0.009319356 which is less than 0.05, Temp variable may not follow
normal distribution.

b) Performing goodness-of-fit on Temp variable by Month variable to check if the variances of mpg are equal or not on am variable categories.
```{r}
# Shapiro-Wilk test for normality by Month
shapiro_test <- airquality %>%
  group_by(Month) %>%
  summarise(p_value = shapiro.test(Temp)$p.value)

# Display the test results
shapiro_test
```
With p-values greater than 0.05, it may follow normal distribution on certain months.
```{r}
# Fit one-way ANOVA to compare Temp by Month
anova_result <- aov(Temp ~ Month, data = airquality)

# Summarize ANOVA results
summary(anova_result)
```

## Question number: 8
Loading the **USArrests** dataset and creating new dataset to work with:
```{r}
data("USArrests")

# Create a new dataset "crime" containing all variables from USArrests
crime <- USArrests

# Print the structure of the new dataset
str(crime)

# Print the first few rows of the new dataset
head(crime)
```

a) Randomly splitting the dataset into training and testing dataset
```{r}
set.seed(20)
index <- sample(2, nrow(crime), replace = T, prob = c(0.8, 0.2))
train.crime = crime[index == 1, ] # Training dataset
test.crime = crime[index == 2, ] # Testing dataset
```

Displaying sample data:
```{r}
head(train.crime)
```



 
Repeating the process for other methods (complete, average, ward)

Fitting Hierarchical Cluster Analysis for other methods
```{r}
hc_complete <- hclust(dist(data[, -1]), method = "complete")
hc_average <- hclust(dist(data[, -1]), method = "average")
hc_ward <- hclust(dist(data[, -1]), method = "ward.D2")
```

Deciding the number of clusters for **k-means** for other methods
```{r}
k_complete <- cutree(hc_complete, k = 4)
k_average <- cutree(hc_average, k = 4)
k_ward <- cutree(hc_ward, k = 4)
```

Fitting **k-means** clustering for other methods
```{r}
km_complete <- kmeans(data[, -1], centers = initial_centers, nstart = 25, algorithm = "Hartigan-Wong")
km_average <- kmeans(data[, -1], centers = initial_centers, nstart = 25, algorithm = "Hartigan-Wong")
km_ward <- kmeans(data[, -1], centers = initial_centers, nstart = 25, algorithm = "Hartigan-Wong")
```

Visualizing the clustering by using **base R graph** and **cluster package** for other methods
Complete Method
```{r}
plot(hc_complete, main = "Complete Linkage", xlab = "", sub = "", cex = 0.6)
rect.hclust(hc_complete, k = 4, border = 2:5)
clusplot(data[, -1], km_complete$cluster, main = "Complete Linkage - k-means", color = TRUE, shade = TRUE)
```
Average Method
```{r}
plot(hc_average, main = "Average Linkage", xlab = "", sub = "", cex = 0.6)
rect.hclust(hc_average, k = 4, border = 2:5)
clusplot(data[, -1], km_average$cluster, main = "Average Linkage - k-means", 
         color = TRUE, shade = TRUE)
```
Ward Method
```{r}
plot(hc_ward, main = "Ward Linkage", xlab = "", sub = "", cex = 0.6)
rect.hclust(hc_ward, k = 4, border = 2:5)
clusplot(data[, -1], km_ward$cluster, main = "Ward Linkage - k-means", 
         color = TRUE, shade = TRUE)
```




```

