---
title: "First Assessment 2080"
author: "13 Kaushal Khatiwada"
date: "2024-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question No: 9

A) 
Column vector X
```{r}
hist(mtcars$mpg)
```

#B)
```{r}
hist(mtcars$mpg,col = "blue",breaks = 10)
```


#C)
```{r}
hist(mtcars$mpg,col = "blue",breaks = 10)
abline(v=mean(mtcars$mpg))
```

#D)

```{r}
#Q-Q plot
qqnorm(mtcars$mpg)
#Q-Q line
qqline(mtcars$mpg,col="red")
```

#E)
```{r}
dens <-density(mtcars$mpg)
#plot density
plot(dens,frame=F,main="Density plot of mpg")
polygon(dens,col="yellow")
```


## Question No: 10
#A)
```{r}
library(igraph)
g <- graph(c(1,2,3,4),directed = F)
```


#B)
```{r}
plot(g)
```


#C)
```{r}
g1 <-graph(c("Sita","Ram","Rita","Gita","Gita","Sita","Sita","Gita","Anita","Rita","Ram","Sita"))
```

#D)
```{r}
plot(g1,
     vertex.color="green",
     vertex.size=20,
     edge.color="red",
     edge.size=10)
```

#E) 
```{r}
degree(g1,mode = "in")
degree(g1,mode = "out")
degree(g1,mode = "all")

```
Degree is number of connections for each node. Sita has the highest degree i.e 4 and Anita  has smallest degree i.e 1 Which means Sita is connected with highest number for vertices. here, mode identities the incoming or outgoing)

```{r}
closeness(g1)

```
Sita is closest and Anita is farthest because Sita has highest number of degree associated with vertices. 

```{r}
betweenness(g1)
```
No vertex will travel through Anita to get to the other vertices
