---
title: "Project 4 Part I"
author: "Kaushal Khatiwada"
date: "2024-05-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1
Using “airquality” Dataset

## A)Get mean and standard deviation of Wind variable by Month variable using the appropriate “apply” family of function, show both the results in a single table and interpret them carefully 	
```{r}
aq <- airquality
aq_means<- tapply(aq$Wind,aq$Month,mean) #Mean
aq_sd <- tapply(aq$Wind,aq$Month,sd)     #Standard Deviation
aq_table <- data.frame(Month = names(aq_means), 
                            Mean_of_Wind = as.numeric(aq_means), 
                            SD_of_Wind = as.numeric(aq_sd))
aq_table
```
## B) Perform goodness-of-fit test on Wind variable by Month variable to check if it follows normal distribution or not
```{r}
library(stats)
library(dplyr)
```

Find the Sample size
```{r}
permonth <- aq %>% group_by(Month) %>% summarize(count= n())
permonth
```

Since, the sample size for each month is <100 we use Shapiro-Wilk normality test
```{r}
tapply(aq$Wind, aq$Month, shapiro.test)
# p-value = 0.4659 , Normally distributed
# p-value = 0.501 , Normally distributed
# p-value = 0.1564 , Normally distributed
# p-value = 0.937 , Normally distributed
# p-value = 0.7852 , Normally distributed
```
p-value of all months is > 0.05  so It follows normal distribution

## C) Perform goodness-of-fit test on Wind variable by Month variable to check if the variances of Wind are equal or not on a variable categories
```{r warning=FALSE}
library(car)
```

Levene Test is performed for variance check because there are >2 categorical group in Month
```{r}
aq$Month <- factor(aq$Month)
leveneTest(aq$Wind ~ aq$Month, data=aq,center=mean)  
```
p-value = 0.9454 i.e > 0.05 so,equal variance among months


## D) Discuss which one-way ANOVA must be used to compare “Wind” variable by “Month” variable categories based on the results obtained above

## ASSUMPTIONS
Normality Test using shapiro-wilk also fulfilled
Variance Test using LeveneTest also fulfilled
Now ready to fit one-way ANOVA


## E) Fit the best one-way ANOVA for this data now and interpret the results carefully
```{r}
summary(aov(aq$Wind ~ aq$Month, data = aq))
```
p-value is 0.008 i.e <0.05. so post-hoc test is required

## F) Fit the most-appropriate post-hoc test if the ANOVA is statistically significant and interpret the result carefully
```{r}
TukeyHSD(aov(Wind ~ Month,data=aq))
```
