---
title: "Project 4 Part III"
author: "Kaushal Khatiwada"
date: "2024-05-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 3
Using "mtcars" Dataset

## A) Divide the mtcars data into train and test datasets with 80:20 random splits
```{r}
set.seed(13)
index <- sample(2,nrow(mtcars),replace = T,prob = c(0.8,0.2)) #Random sampling into two independent c=variable
train.mtcars <- mtcars[index==1,] #Training set
test.mtcars <- mtcars[index==2,]  #Test set
```


## B) Fit a supervised logistic regression model classification models on train data with “am” as dependent variable and all other variables as independent variable

```{r warning=FALSE}
train.suplogr <- glm(am ~., data=train.mtcars,family=binomial)

```


## C) Check multicollinearity of this model and finalize it using appropriate VIF cut-off value for logistic regression 
```{r}
library(car)
vif(train.suplogr)
```
Remove "Disp" because Independent variables with VIF >2 means presence of multicollinearity.
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ mpg+cyl+hp+drat+wt+qsec+vs+gear+carb, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Remove "Qsec" because Independent variables with VIF >2 means presence of multicollinearity.
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ mpg+cyl+hp+drat+wt+vs+gear+carb, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Remove "Cyl" because Independent variables with VIF >2 means presence of multicollinearity
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ mpg+hp+drat+wt+vs+gear+carb, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Remove "Carb" because Independent variables with VIF >2 means presence of multicollinearity
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ mpg+hp+drat+wt+vs+gear, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```

Remove "Mpg" because Independent variables with VIF >2 means presence of multicollinearity
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ hp+drat+wt+vs+gear, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Remove "Hp" because Independent variables with VIF >2 means presence of multicollinearity
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ drat+wt+vs+gear, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Remove "Gear" because Independent variables with VIF >2 means presence of multicollinearity
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ drat+wt+vs, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Remove "Vs" because Independent variables with VIF >2 means presence of multicollinearity
```{r warning=FALSE}
train.suplogr1 <- glm(am ~ drat+wt, data=train.mtcars,family=binomial)
vif(train.suplogr1)
```
Now, VIF<2 so, "Drat" and "wt" remains




