---
title: "Question_No_10"
author: "Utsab Bhattarai"
date: "2024-05-31"
output:
  pdf_document: default
  html_document: default
---

# Use the first four variables of the "iris" data and do as follows with R script to knit PDF output:
```{r}
data("iris")
```

## a. Fit a hierarchical clustering model using single linkage and get the dendogram for this model.
```{r}
iris_data <- iris[, 1:4]
dist_iris <- dist(iris_data)
hc_single <- hclust(dist_iris,
                    method = "single")
plot(hc_single,
     main = "Dendrogram for Single Linkage",
     sub = "",
     xlab = "",
     cex = 0.9)
```

## b. Fit a hierarchical clustering model using complete linkage and get the dendogram for this model.
```{r}
hc_complete <- hclust(dist_iris, method = "complete")
plot(hc_complete,
     main = "Dendrogram for Complete Linkage",
     sub = "",
     xlab = "",
     cex = 0.9)
```

## c. Fit a hierarchical clustering model using average linkage and get the dendogram for this model.
```{r}
hc_average <- hclust(dist_iris,
                     method = "average")
plot(hc_average,
     main = "Dendrogram for Average Linkage",
     sub = "",
     xlab = "",
     cex = 0.9)
```

## d. Find the best hierarchical clustering model, model for this data and locate the number of clusters for it.
```{r}
library(cluster)
sil_width_single <- silhouette(cutree(hc_single,
                                      k = 3),
                               dist_iris)
sil_width_complete <- silhouette(cutree(hc_complete,
                                        k = 3),
                                 dist_iris)
sil_width_average <- silhouette(cutree(hc_average,
                                       k = 3),
                                dist_iris)
# Calculate average silhouette width for each model
avg_sil_single <- mean(sil_width_single[,
                                        3])
avg_sil_complete <- mean(sil_width_complete[,
                                            3])
avg_sil_average <- mean(sil_width_average[,
                                          3])
cat("Average Silhouette Width for Single Linkage:", avg_sil_single, "\n")
cat("Average Silhouette Width for Complete Linkage:", avg_sil_complete, "\n")
cat("Average Silhouette Width for Average Linkage:", avg_sil_average, "\n")
best_model <- which.max(c(avg_sil_single,
                          avg_sil_complete,
                          avg_sil_average))
model_names <- c("Single Linkage",
                 "Complete Linkage",
                 "Average Linkage")
cat("The best hierarchical clustering model is:", model_names[best_model], "\n")
best_hc <- switch(best_model,
                  hc_single,
                  hc_complete,
                  hc_average)
```