---
title: "Question_No_7"
author: "Utsab Bhattarai"
date: "2024-05-31"
output:
  pdf_document: default
  html_document: default
---

# Do the followings in R studio using "airquality" data set of R script to knit PDF output:
```{r}
data("airquality")
```

## a. Perform goodness-of-fit test on Temp variable to check it it follows normal distribution or not.
```{r}
library(stats)
shapiro_test <- shapiro.test(airquality$Temp)
shapiro_test
```

Interpretation:
The Shapiro-Wilk test for normality checks if the Temp variable follows a normal distribution.
Here, the p-value (0.009319) is less than 0.05, thus we reject the null hypothesis that the data follows a normal distribution.

## b. Perform goodness-of-fit test on Temp variable by Month variable to check if the variances of mpg are equal or not on am variable categories.
```{r}
bartlett_test <- bartlett.test(Temp ~ factor(Month), data = airquality)
bartlett_test
```

Interpretation:
Bartlett Test checks if the variances of Temp are equal across the different months.
Here, the p-value (0.01718) is less than 0.05, we reject the null hypothesis that the variances are equal.

## c. Discuss which independent sample must be used to compare "Temp" variable by "Month" variable categories based on the results obtained above.

Interpretation:
As the Temp data follows a normal distribution and the variances are equal, we use ANOVA.

## d. Perform the best independent sample statistical test for this data now and interpret the results carefully.
```{r}
anova_test <- aov(Temp ~ factor(Month), data = airquality)
summary(anova_test)
```

Interpretation:
ANOVA checks if there are any statistically significant differences in the means of Temp across the different months.
Here, the p-value is less than 0.05, we reject the null hypothesis that the means are equal.