---
title: "Assignment 2 (Session 7, 8 & 9)"
author: "Kaushal Khatiwada"
date: "2024-03-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading JSON file from URL
JSON (JavaScript Object Notation) is a text-based format for storing and transporting data.
```{r warning=FALSE}
library(jsonlite)
Raw<- fromJSON("https://data.ny.gov/api/views/9a8c-vfzj/rows.json?accessType=DOWNLOAD")
food_market<-Raw[['data']]    #Fetch data
Names<-food_market[,14]     #Get 14th Column only
head(Names)
```

Representation in Table format with frequency
```{r}
head(table(food_market[,19]))
```
## HTML Scrapping with R
We can get the information from the internet using web scrapping but might be illegal if done without authorization 
```{r warning=FALSE}
library(rvest)
simple<-read_html("https://dataquestio.github.io/web-scraping-pages/simple.html")
simple %>% 
  html_nodes("p") %>%   #Get <p> HTML tag
  html_text()           #Extract text content
```

<h3>Scrape Covid Table from wikipedia</h3>
Scrapping using html elements.
```{r}
library(rvest)
wiki_link<-"https://en.wikipedia.org/wiki/COVID-19_pandemic_in_Nepal"
wiki_page<-read_html(wiki_link)
covid_table<-wiki_page %>% 
  html_elements(".COVID-19_pandemic_data_Nepal_medical_cases") %>%    #Div class
  html_nodes("table") %>%                                     #Get table tag
  html_table() %>%.[[1]]                                      #Get table content
tail(covid_table)
```
<h2>Data Wrangling</h2>

Previously, Covid Table contains sub column [Total,New,Active]. so, we need to remove the sub column heading after concatenating the column with sub column using "_" separator.
```{r}
names(covid_table)<- paste(names(covid_table),covid_table[1,], sep="_")
covid_table<-covid_table[-1,]
tail(covid_table)
```

<h3>Renaming Column names</h3>
```{r}
colnames(covid_table) <- c("Date", "Confirmed_Cases_Total", 
                           "Confirmed_Cases_New", "Confirmed_Cases_Active", 
                           "Recoveries_Total", "Recoveries_New", "Deaths_Total", 
                           "Deaths_New", "PCR_Total", "PCR_New", "TPR", "RR", "CFR", "Ref")
tail(covid_table)
```

<h3>Remove "+" , "%" and ","</h3>
Cleaning the data using gsub()function to replace the pattern.<br>
gsub(PATTERN,REPLACEMENT,COLUMN)
```{r}
covid_table$Confirmed_Cases_New <- gsub('[+]', '', covid_table$Confirmed_Cases_New)
covid_table$Recoveries_New <- gsub('[+]', '', covid_table$Recoveries_New)
covid_table$Deaths_New <- gsub('[+]', '', covid_table$Deaths_New)
covid_table$PCR_New <- gsub('[+]', '', covid_table$PCR_New)

covid_table$TPR <- gsub('[%]', '', covid_table$TPR)
covid_table$RR <- gsub('[%]', '', covid_table$RR)
covid_table$CFR <- gsub('[%]', '', covid_table$CFR)

covid_table$Confirmed_Cases_Total <- gsub('[,]', '', covid_table$Confirmed_Cases_Total)   #Remove "," from "Confirmed_Cases_Total" column's record
covid_table$Confirmed_Cases_New <- gsub('[,]', '', covid_table$Confirmed_Cases_New)
covid_table$Confirmed_Cases_Active <- gsub('[,]', '', covid_table$Confirmed_Cases_Active)
covid_table$Recoveries_Total <- gsub('[,]', '', covid_table$Recoveries_Total)
covid_table$Recoveries_New <- gsub('[,]', '', covid_table$Recoveries_New)
covid_table$Deaths_Total <- gsub('[,]', '', covid_table$Deaths_Total)
covid_table$Deaths_New <- gsub('[,]', '', covid_table$Deaths_New)
covid_table$PCR_Total <- gsub('[,]', '', covid_table$PCR_Total)
covid_table$PCR_New <- gsub('[,]', '', covid_table$PCR_New)

tail(covid_table)
```

<h3>Convert data type from characters to integers and numeric
```{r}
covid_table$Confirmed_Cases_Total <- suppressWarnings(as.integer(covid_table$Confirmed_Cases_Total)) 
covid_table$Confirmed_Cases_New <- suppressWarnings(as.integer(covid_table$Confirmed_Cases_New))
covid_table$Confirmed_Cases_Active <- suppressWarnings(as.integer(covid_table$Confirmed_Cases_Active))
covid_table$Recoveries_Total <- suppressWarnings(as.integer(covid_table$Recoveries_Total))
covid_table$Recoveries_New <- suppressWarnings(as.integer(covid_table$Recoveries_New))
covid_table$Deaths_Total <- suppressWarnings(as.integer(covid_table$Deaths_Total))
covid_table$Deaths_New <- suppressWarnings(as.integer(covid_table$Deaths_New))
covid_table$PCR_Total <- suppressWarnings(as.integer(covid_table$PCR_Total))
covid_table$PCR_New <- suppressWarnings(as.integer(covid_table$PCR_New))

covid_table$TPR <- as.numeric(covid_table$TPR)
covid_table$RR <- as.numeric(covid_table$RR)
covid_table$CFR <- as.numeric(covid_table$CFR)
str(covid_table)
```
<h3>Convert Date Format</h3>
Date was in the format of %d %b. So it is converted to %m %d.
```{r}
date_format <- as.Date(covid_table$Date,format = "%d %b")
final_format <- format(date_format,"%m-%d") 
covid_table$Date <- final_format
tail(covid_table)
```
## Tibble
A tibble is a type of data frame.
<h3>Create Tibble</h3>
```{r}
library(tidyr)
table1 <- tibble(
  country = c("Afghanistan","Afghanistan","Brazil","Brazil","China","China"),
  year = c(1999,2000,1999,2000,1999,2000),
  cases = c(745,2666,37737,80488,212258,213766),
  population = c(19987071,20595360,172006362, 174504898, 1272915272,1280428583)
)
table1
```

<h3>Pivoting [Longer to Wider]</h3>
Best for standard statistical analysis
```{r}
table2
table2 %>%pivot_wider(names_from = type, values_from = count)
```
<h3>Pivoting [Wider to Longer]</h3>
Best for variance components analysis
```{r}
table4a
table4a %>% pivot_longer(c('1999','2000'),names_to = "year",values_to="cases")
```
<h3>Separate</h3>
Separate the single column into multiple column
```{r}
table3
table3 %>% separate(rate,into = c("cases","population"))
```
<h3>Unite</h3>
Combine multiple columns into single column
```{r}
table5
table5 %>% unite(new,century,year)
```

<h4>Missing values in tibble</h4>
Explicitly missing: 4th Quarter of 2015<br>
Implicitly missing: 1st Quarter of 2016
```{r}
stocks <- tibble(
  year = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr = c( 1, 2, 3, 4, 2, 3, 4),                          
  return = c(1.88, 0.59, 0.35, NA, 0.92, 0.17, 2.66)
)
stocks

```

<h4>Pivot and changing the data set to handle the missing values</h4>
```{r}
stocks
stocks %>% pivot_wider(names_from = year,values_from = return)
```

<h4>Handle the missing value and drop the NA value.
```{r}
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(
    cols = c(`2015`, `2016`), 
    names_to = "year", 
    values_to = "return", 
    values_drop_na = TRUE
  )
```

<h4>Using "complete" function to handle missing value
```{r}
stocks %>% complete(year,qtr)
```

<h4>Tibble by row </h4>
```{r}
treatment <- tribble(
  ~ person, ~ treatment, ~response,
  "Derrick Whitmore", 1, 7,
  NA, 2, 10,
  NA, 3, 9,
  "Katherine Burke", 1, 4
)
treatment
```

<h4>Use fill function to handle missing values</h4>
```{r}
treatment %>% fill(person)  
```


## Data Manipulation
filter() function is used to pick the observation by their values.<br>
arrange() function is used to reorder the rows.<br>
select() function is used to pick the variables by their names.<br>
mutate() function is used to create new variable with respect to existing variables<br>
summarise() function is used to merge down the values to single summary.<br>
Use the NYC flight on year 2013 data set.<br>
<br>
<br>
<h3>Filter</h3>
Filter flights for 25th December 2013
```{r warning=FALSE}
library(dplyr)
library(nycflights13)                 #nycflights13 data
filter(flights, month == 12,day == 25)  #filter fights on Christmas day
```
Using Operator to filter the flight on Either November or December
```{r}
filter(flights,month==11 | month==12) 
#          OR
#filter(flights,month==11 | 12)
#filter(flights,month %in% c(11,12))
```
De Morgan's Law

To filter the arrival delay and departure delay less than 120 minutes.
```{r}
filter(flights,!(arr_delay>120 | dep_delay>120))
#         OR
#filter(flights,arr_delay<=120 , dep_delay<=120)
```

<h3>Arrange</h3>
To reorder the data set in year,month,day in descending order 
```{r}
arrange(flights,year,month,day)     
```
Most delayed Airlines<br>
      arrange departure delay in descending order<br>
      Get Top 20<br>
      Represent the carrier and its dep_delay frequency in a table<br>
```{r}
delay<-arrange(flights,desc(dep_delay))  
dep_delay_20<- head(delay,n=20)                           
most_delayed_airlines_table<-table(dep_delay_20$carrier)  
most_delayed_airlines_table
```
Average departed delayed time 
```{r}
average_departed_delay<-mean(dep_delay_20$dep_delay)
average_departed_delay
```

<h3>Select</h3>
```{r}
select(flights,year,month:day)    #Column between month and day
select(flights,-(year:day))     #Exclude column between year and day
select(flights,time_hour,air_time)
```

<h3>Mutate</h3>
Adds new column to existing column<br>
Mutate gain,speed,hours and gain_per_hour column in existing.
```{r}
flights_sml<-select(flights,
                    year:day,
                    ends_with("delay"),
                    distance,
                    air_time
)
flights_sml
mutate(flights_sml,
       gain=dep_delay - arr_delay,
       speed= distance / air_time * 60,
       hours = air_time /60,
       gain_per_hour = gain /hours
)
```
<h3>Transmute</h3>
It is used only to keep the new variables 
```{r}
transmute(flights,
          gain=dep_delay - arr_delay,
          hours = air_time/60,
          gain_per_hour=gain/hours,
)

transmute(flights,
          dep_time,
          hour =dep_time %/% 100,      #Modular
          minute =dep_time %% 100,     #Remainder
)
```
<h3>Summarise</h3>
To group the variable by value
```{r}
summarise(flights,delay=mean(dep_delay,na.rm=TRUE))
by_day<-group_by(flights,year,month,day)
summarise(by_day,delay=mean(dep_delay,na.rm=TRUE), .groups = 'drop')
```

## MULTIPLE OPERATION WITH PIPES
Displays the mean distance of flights and mean arrival delay of flights to each destination with more than 20 flights excluding destination "HNL"
```{r}
delays<-flights %>% 
  group_by(dest) %>% 
  summarise(
    count=n(),
    dist=mean(distance,na.rm = TRUE),
    delay = mean(arr_delay,na.rm = TRUE),
  ) %>% 
  filter(count>20,dest != "HNL")
delays
```
Remove Cancelled flights
  First, Filters only those rows where departure delay and arrival delay is not missing which means the flight has take-off and landed.
```{r}
not_cancelled<- flights %>% 
  filter(!is.na(dep_delay),!is.na(arr_delay))
not_cancelled
```
  Secondly, Calculate the mean departure delay for each day.
```{r}
not_cancelled %>% 
  group_by(year,month,day) %>% 
  summarise(mean=mean(dep_delay), .groups = 'drop') 

``` 
```{r}
  hist(delays$delay)
```


## Summary Function
(A) When do the first and last flights leave each day?<br>
Calculate the minimum departure time and maximum departure time every day using min and max value of departure time
```{r}
not_cancelled %>% 
  group_by(year,month,day) %>% 
  summarise(
    first= min(dep_time),
    last = max(dep_time), .groups = 'drop'
  )
```
On Jan 1 first departure time is 5:17 AM and last departure time is 11:56 PM.


(B)  Why is distance to some destinations more variable than to others?<br>
Calculates the mean distance and standard deviation distance(in descending order)
```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(
    distance_mean=mean(distance),distance_sd=sd(distance)) %>% 
  arrange(desc(distance_sd))
```


(C)  Which destination have the most carries?

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers=n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```
"ATL","BOS","CLT","ORD","TPA" has the most carries


(D) How many flights left before 5am daily?<br>
Shows the number of flights departing earlier than 5:00 AM everyday
```{r}
not_cancelled %>% 
  group_by(year,month,day) %>% 
  summarise(n_early=sum(dep_time<500), .groups = 'drop')
```
On Jan 1 no flight has departed earlier than 5:00 AM


(E) what proportion of flight are delayed by more than an hour?
```{r}
not_cancelled %>% 
  group_by(year,month,day) %>% 
  summarise(hour_prop=mean(arr_delay>60),.groups = 'drop')
```

(F) Find all groups bigger than a threshold?<br>
Display all the flights only to the destination that have more than flights in total.
```{r}
popular_dests<- flights %>% 
  group_by(dest) %>% 
  filter(n()>365)
popular_dests
```



(G) Popular destination
Group By destination and get the total flights and arrange in descending order.
```{r}
popular_destination <- flights %>%
  group_by(dest) %>%
  summarise(total_flights = n()) %>%
  filter(total_flights > 365) %>%
  arrange(desc(total_flights))
head(popular_destination$dest)
```
"ORD" is the popular destination


<h3>Slice function</h3>
```{r}
flights %>% slice_sample(n=5,replace = TRUE)
set.seed(123)
train_data <- flights %>% slice_sample(prop = 0.8)
train_data
test_data <- flights %>% slice_sample(prop = 0.2)
test_data
```

