---
title: "Untitled"
author: "Prashant Jha"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(stats)

library(car)

library(ggcorrplot)

library(lmtest)

library(caret)

```
##question 7
```{r}

library(tidyverse)

aq <- airquality

wind_by_month <- sapply(split(aq, aq$Month), FUN = function(x) c(mean(x$Wind), sd(x$Wind)))

wind_by_month_df <- as.data.frame(t(wind_by_month))

colnames(wind_by_month_df) <- c('Mean', 'SD')

print(wind_by_month_df)

```

#### This table is showing the Mean and Standard Deviation of the speed of wind of everyday from month May to September (represented by 5 to 9) in the year 1973.
 
 
#### a.

#### I am using Shapiro-Wilk test to perform goodness-of-fit test on wind variable to check if it follows normal distribution or not.

```{r}

library(stats)

hist(aq$Wind)

shapiro.test(aq$Wind)

```

#### Since the W value is closer to 1 and p-value is >0.05, I conclude that the variable Wind is normally distributed.
 
 
#### b.

#### Goodness-of-fit test to check if variances among different groups(months) are equal or not.

```{r}

library(car)

aq$Month <- factor(aq$Month)

leveneTest(aq$Wind ~ aq$Month, data = aq, center = mean)

```

#### Since p-value > 0.05, we accept the null hypothesis that means equal variance among months.
 
#### c.

#### Normality test of each month.

```{r}

with(aq, shapiro.test(Wind[Month == 5]))

with(aq, shapiro.test(Wind[Month == 6]))

with(aq, shapiro.test(Wind[Month == 7]))

with(aq, shapiro.test(Wind[Month == 8]))

with(aq, shapiro.test(Wind[Month == 9]))

```

#### Since the W value is closer to 1 and p-value is >0.05 in each month, I conclude that variable Wind in each Month is normally distributed.So classical one-way anova must be used.
 
 ```{r}
 dataset <- iris
dataset
 ```
 
## 10.a),b),c)
## Putting the iris data set and creating a hierarchical clustering using these methods.
 ```{r}
 hcasingle <- hclust(dist(dataset),method = "single")
hcacomplete <- hclust(dist(dataset),method = "complete")
hcaaverage <- hclust(dist(dataset),method = "average")
 ```
 
#### Getting the dendrogram using three different models i.e. single linkage, complete linkage,
## average linkage
 
 ```{r}
 plot(hcasingle, main = "Single Linkage",xlab="",sub="",cex = 0.6)
plot(hcacomplete, main = "Complete Linkage",xlab="",sub="",cex = 0.6)
plot(hcaaverage, main = "Average Linkage",xlab="",sub="",cex = 0.6)
 ```
 
 ```{r}
 
 ```
 
 ```{r}
 
 ```
 