---
title: "Project 2 Part 2"
author: "Kaushal Khatiwada"
date: "2024-03-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web scraping and data wrangling
```{r warning=FALSE}
library(rvest)
library(jsonlite)
```

# https://data.covid19india.org (two JSON files)

Find all the hyperlinks in the site and grep the .json link
```{r}
covid_india_urls <- read_html("https://data.covid19india.org") %>% 
  html_elements("a") %>% 
  html_attr("href") 

# Filter URLs with ".json" extension
urls <- covid_india_urls[grep("\\json$",covid_india_urls)]
urls
```
Extract all the data set from json url.
```{r}
covid_india_timeseries <- lapply(urls[1],fromJSON)
timeseries <- covid_india_timeseries[[1]]

```

```{r}
covid_india_data <- lapply(urls[2],fromJSON)
datas <- covid_india_data[[1]]

```

## Datawrangling on json

Data set is in the nested list. Tried to clean and format the data set.

```{r}
# For fist json link
# Tried to get values
timeseries$WB$dates$`2021-10-31`$total$vaccinated1   #56192166

#Converted to data Frame
timeseries_df <- as.data.frame(timeseries)

# Unlisted all the elements from the list
test <- unlist(timeseries, recursive = T)

```
```{r}
head(test)
```

```{r}
# For fist json link
# Tried to get values
datas$AN$districts$`South Andaman`$total$vaccinated1    #189662
datas$WB$districts$Darjeeling$total$vaccinated1         #1324555


data_df <- as.data.frame(covid_india_data)
```
## AQI FORECAST TABLE
From https://aqicn.org/forecast/kathmandu (aqi forecast table)
```{r}
#aqi_forecast_table <- read_html("https://aqicn.org/city/kathmandu") %>%
#  html_nodes("table") %>%
#  html_table() %>% .[[6]]
# aqi_forecast_table
#head(aqi_forecast_table)

```

Note: Could not get appropriate table from Base R package so used "Rselenium" package for web scrapping

```{r warning=FALSE}
library(RSelenium)
library(netstat)
library(rvest)
```


Scrape the table with table class "aqiforecast-table" by using web client  
```{r}
remote_driver <- rsDriver(remoteServerAddr = "localhost", browser = "firefox", port=free_port(), chromever = NULL, verbose = F)
mybrowser <- remote_driver$client
mybrowser$navigate("https://aqicn.org/forecast/kathmandu/")

aqi_html  <- read_html(mybrowser$getPageSource() %>% unlist())  
aqi_html %>% html_element("table.aqiforecast-table") %>% html_table() -> forecast_table
remote_driver[["server"]]$stop()    #Close the session
```

Data Wrangling to remove the NA values, remove unnecssary rows are columns
```{r}
df <- data.frame(forecast_table)

#Remove all column with NA 
new_df<- df[,colSums(is.na(df))==0]

#Replace column names with first row of data frame
names(new_df) <- new_df[1,]

#Remove 1st row
forecast_df <- new_df[-1,]

#Remove empty row
forecast_df <- forecast_df[!apply(forecast_df == "", 1, all), ]
forecast_df
```

