---
title: "First Reassessment"
author: "Kaushal Khatiwada"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q no 6

A)
Three approach to define integers
```{r}
# Using Vector
r<-c(1:15)
r
```

```{r}
#Using list
l<- list(1:15)
l
```

```{r}
#Using Array
a<-array(1:15)
a
```


B) 
Coulmn vector of size 5
```{r}
N<- c(1.1,2.2,3.3,4.4,5.5)
N
str(N)
```
It is of numeric data type

C) 
Vector Operation
```{r}
add <- r+N
add
```
```{r}
subtract <- N-r
subtract
```

```{r}
multiply <- N*r
multiply
```

```{r}
divide <- N/r
divide
```
Interpret: Since, the size of vector N is 5 and r is 10 the vector operation is repeated even the size is not matched. Vector r repeatedly does operation with N.

D)
List is defined as the data type which can hold the value of same or different data type
```{r}
L<- list("This","is","my","first","programming","in","R")
L
```

## Q no 7
A)
```{r}
aq <- airquality
str(aq)
```
B)
```{r}
```

C)
```{r}
summary(aq)
```

D)
```{r}
library(dplyr)
```
```{r}
temp <- aq$Temp
```
```{r}
data <- aq %>% 
  group_by(aq$Month)
data
```

## Q no 8
A) 
```{r}
barplot(aq$Month)
```

B)
```{r}
hist(aq$Temp)
```
It form a bell-shaped normal distribution.

C)
```{r}
plot(aq$Temp,aq$Day)
```
D)
```{r}
plot(aq$Ozone,aq$Solar.R)
```

## Q no 9
A)
```{r}
country_data <- tibble(
  Country = c("USA", "Nepal", "UK", "Germany", "France", 
              "Italy", "Spain", "Japan", "China", "Australia",
              "USA", "Nepal", "UK", "Germany", "France", 
              "Italy", "Spain", "Japan", "China", "Australia"),
  Year = c(2020, 2021, 2022, 2020, 2021, 
           2022, 2020, 2021, 2022, 2020,
           2020, 2021, 2022, 2020, 2021, 
           2022, 2020, 2021, 2022, 2020),
  Cases = c(10000, 12000, 15000, 8000, 9500, 
            11000, 7500, 8200, 10500, 6000,
            10000, 12000, 15000, 8000, 9500, 
            11000, 7500, 8200, 10500, 6000),
  Population = c(330000000, 37000000, 67000000, 83000000, 67000000, 
                 60000000, 47000000, 126000000, 1400000000, 25000000,
                 330000000, 37000000, 67000000, 83000000, 67000000, 
                 60000000, 47000000, 126000000, 1400000000, 25000000)
)
country_data
```
B)
```{r}
country_data$LnCase <- log(country_data$Cases)
country_data$LnPop <- log(country_data$Population)
country_data
```

C)
Cases and population
```{r}
plot(country_data$Cases, country_data$Population)
```

LnCase and population
```{r}
plot(country_data$LnCase, country_data$Population)
```

Case and LnPop
```{r}
plot(country_data$Cases, country_data$LnPop)
```

LnCase and LnPop
```{r}
plot(country_data$LnCase, country_data$LnPop)
```
D)
```{r}
par(mfrow=c(2,2))
plot(country_data$Cases, country_data$Population, 
     main = "Cases Vs Population", 
     xlab='Cases',ylab = 'Population',)
plot(country_data$LnCase, country_data$Population,
     main = "LnCases Vs Population", 
     xlab='LnCases',ylab = 'Population')
plot(country_data$Cases, country_data$LnPop, 
     main = "Cases Vs LnPopulation",
     xlab='Cases',ylab = 'LnPopulation')
plot(country_data$LnCase, country_data$LnPop,
     main = "LnCases Vs LnPopulation", 
     xlab='LnCases',ylab = 'LnPopulation')
par(mfrow=c(1,1))
```
## Q no 10
```{r warning=FALSE}
library(igraph)
```
A)
```{r}
g <- graph(c(1,2))
plot(g)
```

B)
```{r}
g1 <- graph(c("S","R","R","G","G","S","S","G","A","R"))
```

C)
```{r}
plot(g1, 
     vertex.color='green',
     vertex.size=30,
     edge.size=5,
     edge.color='red')
```

D)
Degree is number of connections for each node. All "S", "R","G" have same degree i.e 3
```{r}
degree(g1)
```


```{r}
closeness(g1)
```
S is closest to other nodes.

Betweeness
```{r}
betweenness(g1,directed = T)
```







