---
title: "Project 5 Part 4"
author: "Kaushal Khatiwada"
date: "2024-05-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Use “Groceries” data available in the “datasets” package

## A) Load “arules” and “arulesViz” libraries
```{r}
library(arules)
library(arulesViz)
```

## B) Load “Groceries” data, check its structure and interpret it carefully
```{r}
data(Groceries)
str(Groceries)
```

## C) Get Frequent Item frequencies using itemFrequencyPlot function and interpret it carefully
```{r}
itemFrequencyPlot(Groceries, topN=20, type="absolute")
```
Whole milk is Frequenty used item


## D) Set a apriori rule with support = 0.001 and confidence = 0.8 and interpret the output carefully
```{r}
apriori_rule <- apriori(Groceries, 
                        parameter = list(supp = 0.001, 
                                         conf = 0.8,
                                         maxlen=10,
                                         target="rules"))

```

## E) Show the top five rules using inspect and round the results to two digits
```{r}
five_rules<- inspect(apriori_rule[1:5])
```



## F) Sort the rule by confidence in decreasing order
```{r}
sorted_rules <- sort(apriori_rule, by = "confidence", decreasing = TRUE)
inspect(sorted_rules[1:5])
```


## G) Use “whole milk” as target item and show the items in “lhs” with decreasing order of confidence and show the top five rules
```{r}
milk_rule <- apriori(Groceries, 
                       parameter = list(supp = 0.001, 
                                        conf = 0.8),
                                        appearance = list(default="lhs",rhs="whole milk"))
```
```{r}
milk_sorted_rules <- sort(milk_rule, by = "confidence", decreasing = TRUE)
inspect(milk_sorted_rules[1:5])
```

## H) Use “whole milk” as target item and show the items in “rhs” with decreasing order of confidence and show the top five rules
```{r}
milk_rule1 <- apriori(Groceries, 
                     parameter = list(supp = 0.001, 
                                      conf = 0.8),
                     appearance = list(default="rhs"))
```


## I) Write summary and conclusion based on your findings above
