---
title: "Project 5 Part 1"
author: "Kaushal Khatiwada"
date: "2024-05-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Use NCI60 data of ISLR2 package
```{r}
library(ISLR2)
```

## A) Define nci labels (NCI$labs) as nci.labs and nci data (NCI$data) and nic.data
```{r}
nci.labs <- NCI60$labs
nci.data <- NCI60$data
```

## B) Check dimension of nci.data object and interpret it carefully
```{r}
dim(nci.data)   # 64 rows 6830 columns
```

## C) Check first four cancer types using nci.labs object
```{r}
head(nci.labs,4)
```

## D) Fit principal component analysis (PCA) on nci.data with scale = TRUE argument as pr.out object
```{r}
pr.out <- prcomp(nci.data,scale=T)
```

## E) Create a plot showing first three PCA components with three different colors
```{r}
plot(pr.out$x[, 1:3], 
     col = c("red", "green", "blue")[factor(nci.labs)],
     pch = 19, 
     xlab = "PC1", 
     ylab = "PC2")
```

## F) Get summary of pr.out object and interpret it carefully
```{r}
summary(pr.out)
```

## G Plot pro.out object and interpret it carefully
```{r}
plot(pr.out)
```

## H) Create custom scatterplots with principal components in x-axis and proportion variance explained (PVE) in y-axis for the first plot and cumulative PVE in the y-axis for the second plot and interpret them carefully
```{r}
par(mfrow = c(1, 2))
var_expalined <- pr.out$sdev^2 / sum(pr.out$sdev^2)
plot(var_expalined, 
     xlab = "Principal Components", 
     ylab = "PVE",
     col="red")
plot(cumsum(var_expalined), 
     xlab = "Principal Components", 
     ylab = "Cumulative PVE",
     col="blue")
```

## I) Perform PCA with varimax rotation and compare it with the PCA result obtained above
```{r}
library(FactoMineR)
pca.1 <- PCA(nci.data, scale = TRUE, ncp = 3, graph = FALSE)
varimax_rot <- varimax(pca.1$ind$coord[, 1:3])
```

## Compare with the PCA result obtained above
```{r}
par(mfrow = c(1, 2))
plot(pr.out$x[, 1:3], 
     col = c("red", "green", "blue")[factor(nci.labs)], 
     pch = 19, xlab = "PC1",
     ylab = "PC2", 
     main = "Original PCA")

plot(varimax_rot$loadings, 
     col = c("red", "green", "blue")[factor(nci.labs)], 
     pch = 19,
     xlab = "PC1", 
     ylab = "PC2", 
     main = "Varimax Rotated PCA")
```

## J) Write summary of the results and conclusion based on your findings

