---
title: "Retail Store Sales"
author: "Kaushal Khatiwada"
date: "2024-04-19"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling

a)Data Collection 
b)Data Inspection
c)Data Cleaning
      Handling Missing Values 
      Removing Duplicates
      Standardizing Formats
d)Exploratory Data Analysis (EDA)


```{r warning=FALSE}
library(tidyverse)
library(dplyr)
library(tidyr)
```

## A) Data Collection
Import Retail Store sales dataset in a command-seperated file.
```{r}
sales<- read.csv("Sales.csv")
```

## B) Data Inspection
Summary of the structure of a dataset.
```{r}
str(sales)
```

Dimensions of dataset.
```{r}
dim(sales)
```

Display few rows of dataset.
```{r}
head(sales)
```

```{r}
colnames(sales)
```
 1. Invoice id: Computer generated sales slip invoice identification number

 2. Branch: Branch of supercenter (3 branches are available identified by A, B and C).

 3. City: Location of supercenters

 4. Customer type: Members for customers using member card and Normal for without member card.

 5. Gender: Gender type of customer

 6. Product line: General item categorization groups - Electronic accessories, Fashion accessories, Food and beverages, Health and beauty, Home and lifestyle, Sports and travel

 7. Unit price: Price of each product in USD

 8. Quantity: Number of products purchased by customer

 9. Tax: 5% tax fee for customer buying

 10. Total: Total price including tax

 11. Date: Date of purchase (Record available from January 2019 to March 2019)

 12. Time: Purchase time (10am to 9pm)

 13. Payment: Payment Type – Cash, Credit card and Ewallet)

 14. COGS: Cost of goods sold

 15. Gross margin percentage: Gross margin percentage

 16. Gross income: Gross income

 17. Rating: Customer rating on shopping experience (On a scale of 1 to 10)
 
 
```{r}
colSums(is.na(sales))
```


```{r}
sum(duplicated(sales))
```
There is no duplicate records.

## c) Data Cleaning
Handling Missing Values
For Unit Price

```{r}
sales %>% filter(!complete.cases(sales$Unit.price))
```
2 records are NA. So, fill the NA as 
Unit price = (Total - 5% Tax)/Quantity

```{r}
sales<- sales %>% 
  mutate(Unit.price = ifelse(is.na(Unit.price),(Total - Tax.5.)/Quantity,Unit.price))
```


```{r}
colSums(is.na(sales))

```
For Quantity

```{r}
sales %>% 
  filter(!complete.cases(Quantity))
```
16 records are NA. So, fill the NA as 

(Total - 5% Tax)/Unit price

```{r}
sales<- sales %>% 
  mutate(Quantity = ifelse(is.na(Quantity),(Total - Tax.5.)/Unit.price,Quantity))
```

```{r}

colSums(is.na(sales))

```


For 5% Tax

```{r}
sales %>% 
  filter(!complete.cases(Tax.5.))
```
14 records are NA. So, fill the NA as 

Total - (Unit price * Quantity)
```{r}
sales<- sales %>% 
  mutate(Tax.5. = ifelse(is.na(Tax.5.),Total - (Unit.price * Quantity),Tax.5.))
```

```{r}

colSums(is.na(sales))

```
For Total
```{r}
sales %>% 
  filter(!complete.cases(Total))
```
14 records are NA. So, fill the NA as 

(Unit price * Quantity) + 5% Tax
```{r}
sales<- sales %>% 
  mutate(Total = ifelse(is.na(Total),(Unit.price * Quantity) + Tax.5.,Total))
```

```{r}

colSums(is.na(sales))

```
For Gross Margin Percentage
```{r}
sales %>% 
  filter(!complete.cases(gross.margin.percentage))
```
```{r}
table(sales$gross.margin.percentage)
```
Here, Gross Margin Percentage is 4.761904762, So, fill NA as 4.761904762
```{r}
sales<- sales %>% 
  mutate(gross.margin.percentage = ifelse(is.na(gross.margin.percentage),4.761904762,gross.margin.percentage))
```

```{r}

colSums(is.na(sales))

```
## EDA
- Univariate Analysis: Studying one variable at a time
- Bivariate Analysis: Studying two variables at a time
- Multivariate Analysis: Studying multiple variables at a time


## Univariate Analysis
```{r}
branches <- table(sales$Branch)
branches
```
```{r}
barplot(branches,
        main = "Branch",
        xlab = "Branches",
        ylab = "Count",
        col = "blue") 
```  


```{r}
cities <- table(sales$City)
cities
```
```{r}
barplot(cities,
        main = "City",
        xlab = "City",
        ylab = "Count",
        col = "red")  # Bar color
        
```

```{r}
cust_type <- table(sales$Customer.type)
cust_type
```
```{r}
barplot(cust_type,
        main = "Customer Type",
        xlab = "Cutomers",
        ylab = "Count",
        col = "green")
```  


```{r}
gender_type <- table(sales$Gender)
gender_type
```
```{r}
gender_lbl <- paste(gender_type)
pie(gender_type, 
    main = "Gender",
    col= c("red","blue"),
    label=gender_lbl)
legend("topright",names(gender_type),
       cex = 0.8, fill = c("red","blue"))
  

```

```{r}
product_line <- table(sales$Product.line)
product_line
```

```{r}
par(pin = c(4, 2))
barplot(product_line,
        main = "Product line",
        ylab = "Count",
        las= 2,
        col = rainbow(6))

legend("topright",  
       legend = names(product_line), 
       fill = rainbow(6),  
       title = "Product Lines",  
       cex = 0.8,)  
```


```{r}
payment_type <- table(sales$Payment)
payment_type
```
3 records are not categorized under any category. SO, fill with mode categorical value.
```{r}
mode_cat<- names(payment_type)[which.max(payment_type)]
mode_cat
```
Replace the null value with "Ewallet."

```{r}
sales$Payment[sales$Payment == ""] <- mode_cat
```

```{r}
payment_type <- table(sales$Payment)
payment_type
```

```{r}
barplot(payment_type,
        main = "Payment Type",
        ylab = "Count",
        col = rainbow(3))

legend("topright",  
       legend = names(payment_type), 
       fill = rainbow(3),  
       title = "Product Lines",  
       cex = 0.8)  
```

For Invoice ID
```{r}
length(unique(sales$Invoice.ID))
```
9 out of 1000 record have missing Invoice ID.






