---
title: "Word Cloud"
author: "Kaushal Khatiwada"
date: "2024-04-23"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE}
library(rvest)
library(tm)
library(wordcloud)
```

```{r}
urls<- "https://thehimalayantimes.com/opinion/navigating-nepals-digital-frontier-understanding-cybersecurity-in-the-digital-age-ensuring-data-safety-and-the-role-of-ai"
```

```{r}

texts <- read_html(urls) %>% 
  html_nodes("p") %>%   #Get <p> HTML tag
  html_text()   
```

Combine all the text into a single corpus
```{r}
myCorpus <- Corpus(VectorSource(texts))
```


```{r}
inspect(myCorpus[17:20])
```
## Text Cleaning
1) Convert to a lowercase
```{r warning=FALSE}
text <- tm_map(myCorpus, content_transformer(tolower))
inspect(text[17:20])
```

2) Remove Numbers
```{r warning=FALSE}
text <- tm_map(text, removeNumbers)
inspect(text[17:20])
```
3) Remove Punctuation
```{r warning=FALSE}
text <- tm_map(text, removePunctuation)
inspect(text[17:20])
```
4) Remove URL
```{r warning=FALSE}
removeURL <- function(X) gsub("http[^[:space:]]*","",X)
text <- tm_map(text,removeURL)
inspect(text[17:20])
```
5) Remove Stopwords.
```{r warning=FALSE}
text <- tm_map(text, removeWords, stopwords("english"))
inspect(text[17:20])
```
6) Remove Whitespaces.
```{r warning=FALSE}
text <- tm_map(text, stripWhitespace)
inspect(text[17:20])
```

## Stemming and lemmatization
Stemming is a text normalization technique used in natural language processing 
(NLP) that reduces words to their root or base form to enhance language understanding.
Example {remove suffixes} : stem of these three words, "connections", "connected", "connects", is “connect”.

Lemmatization analyzes the contexts of the sentences.
Example: “Saw” would return as “see” or “saw” depending on the context of the word.

```{r}
# Backup the text just in case
backup_text <- text
```

```{r warning=FALSE}
backup_text <- tm_map(backup_text, stemDocument)
inspect(backup_text[17:20])
```
It might not be good idea to use stemming in some cases because "artifici","intellig" is not the real words. So ww will not be using stemmed words.

## Term-Document matrix
Text-Document Matrix represent the text data in the form of a matrix where row correspond to the term in the document and columns correspond to the document in the corpus and cells correspond to the weights of the term. 
```{r}
mytdm <- TermDocumentMatrix(text,control = list(wordLengths = c(1,Inf)))
mymatrix <- as.matrix(mytdm)
inspect(mytdm)
```
Term Frequency represent the frequency of terms in the document.

```{r}
freq.terms <- findFreqTerms(mytdm, lowfreq = 2)
freq.terms
```
To search for term associated with "ai" with correlation coefficient of at least 0.2
```{r}
findAssocs(mytdm, "ai", 0.2)
```

```{r}
freq <- sort(rowSums(mymatrix),decreasing = TRUE)

```

```{r}
library(RColorBrewer)
```

## Word Cloud
```{r warning=FALSE}
set.seed(123)
wordcloud(words = names(freq), freq = freq, min.freq  = 2, random.order = FALSE, colors = brewer.pal(8,"Dark2"))
```