---
title: "Project 3"
author: "Kaushal Khatiwada"
date: "2024-04-28"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WORK 1 
## Locate Median of “Temp” variable graphically

 Find the interval size and class interval using sturge's formula.
```{r}
aq_temp <- airquality$Temp
High <- max(aq_temp)  # max value is 97
Low <- min(aq_temp)  # min value is 56
N = length(aq_temp)  # number of observation is 153

# Calculate number of class intervals Using Sturges' formula
class_intervals <-  ceiling(log2(N) + 1) # 9 class intervals
breaks <- seq(Low,High,length.out = class_intervals)

#Frequency
f <- table(cut(aq_temp,breaks = breaks,include.lowest = TRUE))
f
```
Get frequency for less than ogive
```{r}
less_than_ogive <- cumsum(f)
less_than_ogive
```

Get frequency for more than ogive
```{r}
more_than_ogive <- rev(cumsum(rev(f)))
more_than_ogive
```

Plot less than and more than ogives 
```{r}
plot(breaks[-1],
     less_than_ogive, 
     type = "o",
     col = "blue",
     xlab = "Temperature",
     xlim = c(50,100),
     ylim = c(0,150),
     ylab = "Cumulative Frequency", 
     main = "Ogive for Temperature")
points(breaks[-1], 
       more_than_ogive, 
       type = "o", 
       col = "red",
       ylim = c(0,150),
       xlim = c(55,100))

legend("topright", legend = c("Less than Ogive", "More than Ogive"), col = c("blue", "red"),
       lty = c(1, 1),lwd = 2,cex = 0.6)
```

Draw a line from the intersection breaks (intersection_lower, intersection_upper)
```{r}
intersection_lower <- breaks[which(less_than_ogive >= more_than_ogive)[1]]
intersection_upper <- breaks[which(less_than_ogive >= more_than_ogive)[1]+1]
intersection_lower
intersection_upper

```

```{r}
#Intersection line

```
Find the median in R
```{r}
median <- median(aq_temp)
cat("Median is ",median)
```

## WORK 2
## Locate Mode of “Temp” variable graphically
Histogram of Temp variable in airquality dataset
```{r}
aq_temp <- airquality$Temp
hist(aq_temp, main = "Temperature",col="white",xlab = "Temperature")
h <- hist(aq_temp,plot=F)
```
Get the tallest bar, left bar and right bar
```{r}
highest_freq <- max(h$counts)       #34
highest_freq
```
```{r}
highest_bar <- which.max(h$counts)  # 6th bar
highest_bar
```
```{r}
left_bar <- highest_bar - 1         # 5th bar
left_bar
```
```{r}
right_bar <- highest_bar + 1        # 7th bar
right_bar
```

Draw a diagonal line from upper bound of largest bar  to upper bound of left bar at top of the bar
```{r}

left_upper <- h$breaks[left_bar + 1]
left_bar_count <- h$counts[left_bar]

highest_bar_upper <- h$breaks[highest_bar + 1]
highest_bar_count <- h$counts[highest_bar]
hist(aq_temp, main = "Temperature",col="white",xlab = "Temperature")
segments(left_upper, left_bar_count, highest_bar_upper, highest_bar_count, col = "darkgreen")
```

Draw a diagonal line from lower bound of largest bar  to lower bound of right bar at top of the bar
```{r}
right_lower <- h$breaks[highest_bar] 
right_bar_count <- h$counts[right_bar]

highest_bar_lower <- h$breaks[right_bar] 
highest_bar_count <- h$counts[highest_bar]

hist(aq_temp, main = "Temperature",col="white",xlab = "Temperature")
segments(left_upper, left_bar_count, highest_bar_upper, highest_bar_count, col = "darkgreen")
segments(right_lower, highest_bar_count, highest_bar_lower, right_bar_count, col = "blue")
```
Find the intersction point using point of intersection formula
```{r}
# Calculate slopes (m) and y-intercepts (b) of the two lines
slope1 <- (highest_bar_count - left_bar_count) / (highest_bar_upper - left_upper)
slope2 <- (right_bar_count - highest_bar_count) / (highest_bar_lower - right_lower)

intercept1 <- left_bar_count - slope1 * left_upper
intercept2 <- highest_bar_count - slope2 * right_lower

cat("Slope of grren line",slope1)
cat("\nSlope of blue line",slope2)
cat("\nIntercept of green line",intercept1)
cat("\nIntercept of blue line",intercept2)
```

Using point of intersection formula
```{r}
intersection_x <- (intercept2 - intercept1) / (slope1 - slope2) 
cat("Temperature value at intesection point is ",intersection_x)  # 80.33333
```
So the mode is 80.333 (approx 81)

Draw a line
```{r}
hist(aq_temp, main = "Temperature",col="white",xlab = "Temperature")
segments(left_upper, left_bar_count, highest_bar_upper, highest_bar_count, col = "darkgreen")
segments(right_lower, highest_bar_count, highest_bar_lower, right_bar_count, col = "blue")
abline(v = intersection_x, col = "red", lty = 2)
```
Mode Using R
```{r}
f <- table(aq_temp)
max_freq <- max(f)
mode <- names(f)[f == max_freq]
cat("Mode is", mode)
```

## WORK 3
## Perform social network analysis of first and second variables
```{r}

library(igraph)
```

Get first and column column
```{r}
sna <- read.csv("Kaushal Khatiwada - SNA_School.csv")
s <- sna[, 1:2]
```

```{r}
net <- graph_from_data_frame(s, directed = T)
```

Number of vertices
```{r}
V(net)   # 52 vertices
```
Number of edges
```{r}
E(net)   # 290 Edges
```
Degree of each vertices
```{r}
degree(net)
```
Hisotgram 
```{r}
degree_table <- table(degree(net))
set.seed(13)
hist(degree(net),
     main = "Degree",
     xlab = "Degree",
     ylab = "Frequency",
     col= "red")
```

Network Diagram
```{r}
plot(net,
     vertex.color = "green",
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.8)
```
Network Diagram using Kamada-kawai layout
```{r}
plot(net,
     layout = layout.kamada.kawai,
     main = "Network Diagram (Kamada-Kawai Layout)",
     vertex.color = rainbow(13),
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.8)
```


Hubs is a node with most outer edges
```{r}
hubs <- hub_score(net)$vector
plot(net,
     main = "Hubs",
     vertex.color = rainbow(13),
     vertex.size = hubs * 30,
     edge.arrow.size = 0.1
     )

```
CC and CB are hubs

Authorities is a node with most inner edges
```{r}
authorities <- authority_score(net)$vector
plot(net,
     main = "Authority",
     vertex.color = rainbow(13),
     vertex.size = authorities * 30,
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.8)
```
CA is authority


Community is form with densely connected noede
```{r}
net <- graph_from_data_frame(s, directed = F)
community_cluster <- cluster_edge_betweenness(net)
plot(community_cluster,
     net,
     vetex.size = 10,
     vertex.label.cex = 0.8
)
```
